---
title: "Assignment 2: Time Series Regression"
author: "Joshua Goldberg"
date: "`r format(Sys.time(), '%B, %d %Y')`"
always_allow_html: yes
output:
  pdf_document: default
  github_document: 
editor_options: 
  chunk_output_type: inline
---

```{r Global-options, include=FALSE}
knitr::opts_chunk$set(fig.width=6, fig.asp=0.618, fig.path='Figs/',
                      warning=FALSE, message=FALSE)
```

```{r Preamble, echo=FALSE}
# Enter package in p_load()
# If package is not installed, p_load() will install and load the package
if(!"pacman" %in% rownames(installed.packages())) {
  install.packages("pacman")
  }
pacman::p_load(tidyverse, ggthemes, here, tsibble, lubridate, tseries, rlang, broom)

# Set default ggplot theme to tufte
theme_set(ggthemes::theme_tufte())
```

```{r Copy-files, echo=FALSE, eval=FALSE}
# Enter files to load to project directory in from = "~/Downloads/your_file_name_here"
file.copy(from = "~/Downloads/Assign 2 TS regression.csv", to = here::here(), 
          overwrite = TRUE, recursive = FALSE, 
          copy.mode = TRUE)
```

# Data Description

```{r}
stock_data <- read_csv("Assign 2 TS regression.csv") %>% 
  mutate(date = dmy(date)) %>% 
  gather(key = stock, value = return, -1)

stock_data_ts <- stock_data %>%
  as_tsibble(key = id(stock), index = date)
```


All are daily stock exchange returns.

ISE: Istanbul stock exchange national 100 index

SP: Standard & Poor™s 500 return index

DAX: Stock market return index of Germany

FTSE: Stock market return index of UK

NIKKEI: Stock market return index of Japan

BOVESPA: Stock market return index of Brazil

# Questions

Determine if all the TS are stationary: 

1. qualitatively: the data for each stock all look stationary. $\mu$ and $\sigma^2$ remain constant overtime. Oscillations are offset by each other.

```{r}
stock_data_ts %>% 
  ggplot(aes(date, return, color = stock)) +
  geom_line() +
  scale_x_date(date_breaks = "6 month", date_minor_breaks = "3 month", date_labels = "%m-%y") +
  scale_color_viridis_d(name = "Stock") +
  facet_wrap( ~ stock) +
  labs(x = "Date",
       y = "Return",
       caption = "date format: MM/YY")
```

2. quantitatively: use __ADF__ and __KPSS__ from package tseries.

```{r warning=TRUE}
(stationary_tests <- stock_data_ts %>% 
  nest(-stock) %>% 
  mutate(adf_test = map(data, ~ suppressWarnings(adf.test(.x$return))),
         kpss_test = map(data, ~ suppressWarnings(kpss.test(.x$return))),
         adf_p_value = map_df(adf_test, ~ glance(.x)) %>% pull(p.value),
         kpss_p_value = map_df(kpss_test, ~ glance(.x)) %>% pull(p.value)))

stationary_tests %>% 
  gather(key = key, value = value, -c(1:4)) %>% 
  ggplot(aes(stock, value, fill = key)) +
  geom_col(position = "dodge") +
  geom_hline(yintercept = .05, linetype = 2) +
  annotate("text", -Inf, .0575, label = "Null Rejection Threshold", hjust = 0, vjust = 1) +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_viridis_d(name = "Hypothesis Test", labels = c("ADF", "KPSS")) +
  labs(title = "Determining stationarity with ADF and KPSS",
       x = "Stock",
       y = "P-Value") +
  theme(legend.position = "top")
```

2. Split the data into train and test, keeping only the last 10 rows for test (from date 9-Feb-11). Remember to use only train dataset. 

```{r}
model_data <- stock_data_ts %>% 
  spread(stock, return)

train <- model_data %>% 
  filter(date < "2011-02-09")

test <- model_data %>% 
  anti_join(train, "date")
```

3. Linearly regress ISE against the remaining 5 stock index returns. Determine which coefficients are equal or better than 0.02 (*) level of significance.

```{r}
lm_model <- lm(ISE ~ BOVESPA + DAX + FTSE + NIKKEI + SP, data = train)
summary(lm_model)
```

```{r}
signif_vars <- function(model) {
  model %>% 
    tidy() %>% 
    slice(-1) %>% 
    filter(p.value < .02) %>% pull(term) 
}

signif_vars(lm_model)
```

Significant variables: `r signif_vars(lm_model)`.

4. For the non-significant coefficients, continue to lag by 1 day until all coefficients are significant at 0.01 (*). Use `slide()` function from package __DataCombine__. Remember you will need to lag, so you slideBy = -1 each step. How many lags are needed for each independent variable?

```{r}
# Define shift function to take a dataframe, variable, and shift direction and return a respective dataframe with the new modified variable
shift_var <- function(.data, .var, .shift_by) {
  .var <- enquo(.var)
  shift_direction <- ifelse(.shift_by > 0, "lead", "lag")
  column_name <- paste0(quo_name(.var), "_", shift_direction, abs(.shift_by))
  
  .data %>% 
    mutate(!! column_name := DataCombine::shift(!! .var, shiftBy = .shift_by, reminder = FALSE))
}

lagged_train <- train %>% 
  shift_var(BOVESPA, .shift_by = -1) %>% 
  shift_var(SP, .shift_by = -2)

lagged_train %>% 
  select(date, contains("lag")) %>% 
  head()
```

Two and one lag(s) were needed for `SP` and `BOVESPA`, respectively.

```{r}
lm_model_lag <- lm(ISE ~ BOVESPA_lag1 + DAX + FTSE + NIKKEI + SP_lag2, data = lagged_train)
summary(lm_model_lag)
```

5. Find correlations between ISE and each independent variable. Sum the square of the correlations. How does it compare to R-squared from #4?

```{r}
cor.test(lagged_train$ISE, lagged_train$BOVESPA)

vars <- c("BOVESPA_lag1", "DAX", "NIKKEI", "FTSE", "SP_lag2")

cors <- map(vars, ~ cor.test(lagged_train$ISE, lagged_train[, .x][[1]])) %>% 
   map_dbl("estimate")

sum(cors^2)
```

Sum the square of the correlations is `r sum(cors^2)`.

6. Concept question 1: why do you think the R-squared in #4 is so much less than the sum of square of the correlations? The much higher result compared to $R^2$ is due to collinearity between the independent variables:

```{r}
cor(lagged_train %>% 
      .[complete.cases(.), ] %>% 
      as_tibble() %>% 
      select(-date, -BOVESPA, -SP) %>% 
      as.matrix()) %>% 
  corrplot::corrplot()
```

7. Take the test dataset and perform the same lags from #4 and call `predict()` function using the lm regression object from #4. Why do you need to use the lm function object from #4? Because this is the model we used for the training data with the lagged sequences, which has statistically significant variables.

```{r}
lagged_test <- test %>% 
  shift_var(BOVESPA, .shift_by = -1) %>% 
  shift_var(SP, .shift_by = -2)

test_predictions <- lagged_test[complete.cases(lagged_test), ] %>% 
  mutate(preds = lm_model_lag %>% predict(newdata = lagged_test[complete.cases(lagged_test), ]),
         squared_errors = (preds - ISE)^2,
         rmse = mean(sqrt(squared_errors))) %>% 
  select(date, ISE, preds, squared_errors, rmse)
```

We see that the predictions roughlt follow the trend, but not perfectly. So we have some bias in the model.

```{r}
test_predictions %>% 
  select(-rmse, -squared_errors) %>% 
  gather(key = key, value = value, -date) %>% 
  ggplot(aes(date, value, color = key)) +
  geom_point() +
  geom_line() +
  scale_color_viridis_d()
```


# Concept question 2: what do you find in #1 and why? 

We find that both qualitatively and quantitatively that the time series are stationary. Since both of these methods agree, the conclusion is likely sound. 
