---
title: "Lecture 1"
author: "Joshua Goldberg"
date: "`r format(Sys.time(), '%B, %d %Y')`"
always_allow_html: yes
output:
  pdf_document: default
editor_options: 
  chunk_output_type: inline
---

```{r Global-options, include=FALSE}
knitr::opts_chunk$set(fig.width=6, fig.asp=0.618, fig.path='Figs/',
                      warning=FALSE, message=FALSE)
```

```{r Preamble, echo=FALSE}
# Enter package in p_load()
# If package is not installed, p_load() will install and load the package
if(!"pacman" %in% rownames(installed.packages())) {
  install.packages("pacman")
  }
pacman::p_load(tidyverse, ggthemes, here)

# Set default ggplot theme to tufte
theme_set(ggthemes::theme_tufte())
```

```{r Copy-files, echo=FALSE, eval=FALSE}
# Enter files to load to project directory in from = "~/Downloads/your_file_name_here"
file.copy(from = "~/Downloads/", to = here::here(), 
          overwrite = TRUE, recursive = FALSE, 
          copy.mode = TRUE)
```

# Introduction

What is time series data: data measured at specific time points

$TS \{Y_t\}$

Two types of time series:
 - uniform: time intervals constant
 - non-uniform: time intervals varying
 
 Uniform time series focuses on the forecast.
 
 $t$ is now. $+t$ is a forecast and $-t$ is historical.
 
 __Random variable__ values are defined by a probability distribution.
 
 Two interesting time series:
- __White noise__ is a collection of uncorrelated random variables $\{W_t\}$ with $\mu=0$ and variance=$\sigma w^2$
- __Random walk__ $y_t=y_{t-1}+w_t$

Orders of moment:

1. mean

2. variance

3. skew

4. kurtosis

Anytime you deal with stochastic data the mean is expected. If it's deterministic, then you just do an average.

Variance: $E[(x-\mu)^2]$

Covariance: $E[(x-\mu_x)(y-\mu_y)]$

Auto covariance: $\gamma_{s,t}=E[(x_s-\mu_x)(x_t-\mu_x)]$
$$\gamma_{k}=E[(x_{t+k}-\mu_x)(x_t-\mu_x)]$$

$\gamma_0=E[(x-\mu)^2]$: this is variance at $\gamma=0$

Auto correlation: $\rho_{s,t}= \frac{\gamma_{s,t}}{\sqrt{cov(x_s,x_s) cov(x_tx_t)}}$

# Stationarity

A time series is stationary means that the data generation process is in statistical equilibrium. The statistical properties (orders) of the data do not change. Non-stationary time series has less tools to work with.

__Strict__: probability behavior of every TS collection $\{x_{t1},x_{t2},x_{t3},...x_{tk}\}$ is identical to the time-shifted $\{x_{t1+h},x_{t2+h},x_{t3+h},...x_{tk+h}\}$

Note: the joint probability distribution of the time series is constant over time.

__Weak__: 

1. $\{y_t\}$ TS finite variance process that has a constant mean that does not change with time.

2. Auto convariance that only depends on the lag $k$.

Weak + Gaussian distribution = strict.

Gaussian distribution is both weak and strict.

If you have a process that is everything __strict__, but is not finite variance, then it is not weak. 

Cauchy example, which has no secord order moments.

Everything in class is weak stationary.

## I.I.D.

iid => strict
  distributions that are independent
  
But strict != iid due to auto correlation.

## White noise

Weak since variance is dependent on lag

## Random walk

Non-stationary because auto covariance changes?

# EDA

If a time series has some auto correlation, you should be able to write:

$y_t=\rho y_{t-1}+\mu_t$, for $\rho<1$

If $\rho=1$, this becomes a random walk (non-stationary)

## Tests

Augmented Dickey Fuller (ADF) test is a hypothesis testing that says $H_0=\rho=0$: non-stationary and $H_a=\rho<0$: stationary.

Kwiat Kowski, Phillips Schmidt, Shin (KPSS) test for stationarity.
$H_0=$ stationary
$H_a=$ non-stationary

KPSS has been proven to be more stable.

```{r}
library(tseries)
```

```{r}
wn <- rnorm(500, 0, 1)
rw <- cumsum(wn)

ts.plot(wn)
ts.plot(rw)
```

```{r}
adf.test(wn)
```

```{r}
kpss.test(wn)
```

```{r}
adf.test(rw)
```

```{r}
kpss.test(rw)
```

When you see an acf that dies down.
```{r}
acf(wn)
```

Conversely, if you see an acf that refuses to die down, that is non-stationary.
```{r}
acf(rw)
```

```{r}
rw <- c(0)
wn <- rnorm(500, 0, 1)
rw[1] <- wn[1]

for (i in 2:length(wn)) {
  rw[i] <- -1 * rw[i-1] + wn[i]
}

ts.plot(rw)
adf.test(rw)
kpss.test(rw)
```

You cannot randomly sample time series because you end up breaking the auto correlation.




















