---
title: "Lecture 3"
author: "Joshua Goldberg"
date: "`r format(Sys.time(), '%B, %d %Y')`"
always_allow_html: yes
output:
  pdf_document: default
editor_options: 
  chunk_output_type: inline
---

```{r Global-options, include=FALSE}
knitr::opts_chunk$set(fig.width=6, fig.asp=0.618, fig.path='Figs/',
                      warning=FALSE, message=FALSE)
```

```{r Preamble, echo=FALSE}
# Enter package in p_load()
# If package is not installed, p_load() will install and load the package
if(!"pacman" %in% rownames(installed.packages())) {
  install.packages("pacman")
  }
pacman::p_load(tidyverse, ggthemes, here, tseries, tsibble, forecast)

# Set default ggplot theme to tufte
theme_set(ggthemes::theme_tufte())
```

```{r Copy-files, echo=FALSE, eval=FALSE}
# Enter files to load to project directory in from = "~/Downloads/your_file_name_here"
file.copy(from = "~/Downloads/", to = here::here(), 
          overwrite = TRUE, recursive = FALSE, 
          copy.mode = TRUE)
```

# Data transformations

1. Decouple mean and variance as close as possible to making them constant

2. Model additive: simpler than exponential

3. Reshape residuals

## Box Cox Transformation

$y_t$ 

$\lambda=0$: $logy_t$

For the above, you must add $x$ to all observations, which will just shift the distribuition

Otherwise: $\frac{y_t^\lambda-1}{x}$

# Simple operators

Backward shift operator $\beta$

$\beta z_t=t_{t-1}$

$\beta^m z_t=t_{t-m}$

Forward shift operator  $F$

$Fz_t=z_{t+1}$

$F^mz_t=z_{t+m}$

$$
\begin{aligned}
\triangledown y_t&=y_t-y_{t-1} \\
&=y_t-\beta y_t \\
&=(1-\beta)y_t
\end{aligned}
$$

## Autoregressive (AR)

Current value is a finite linear regression of past values:

Define: $\Phi(\beta)=1-\phi\beta-\phi_2\beta^-...-\phi_p\beta^p$

$$
\begin{aligned}
y_t&=\phi_1y_t-1+\phi_2y_t-2+...+\phi_py_t-p+\epsilon_t \\
&=\phi\beta y_t+\phi_2\beta^2 y_t+...+\phi_p\beta^p y_t+\epsilon_t \\
&=y_t-\phi_1\beta y_t-\phi_2\beta^2 y_t-...-\phi_p\beta^p y_t = \epsilon_t \\
(1-\phi_1\beta-\phi_2]beta^2-...-\phi_p\beta^p) &= \epsilon_t
\end{aligned}
$$
Finally:

$\Phi(\beta)y_t=\epsilon_t$ is an auto-regressive process.

When you find a characteristic function of the data generation process, you want find out if it has a unit-root; if it does, it is guranteed to be non-stationary. If any root relies on unity, we know the time series cannot be stationary; an autogressive time model only works with stationary time series.

Linear regression only works with stationary time series.

Very few natural processes have an instantenous effect. You need to figure out the lag effect.

$$
AR(1) \\
[y_t=\phi_1y_{t-1}+\epsilon_t]y_{t-k} \\
E[y_ty_{t-k}=\phi_1y_{t-1}y_{t-k}+\epsilon_ty_{t-k}]
$$

$\phi<1=$ stationary time series.

Always: $\epsilon_ty_{t-k}=0$

If $\phi$ is negative, acf will oscillate. 

## Moving average (MA)

$y_t=\epsilon_t-\theta_1\epsilon_{t-1}-\theta_2\epsilon_{t-2}-...-\theta_q\epsilon_{t-q}$

Define: $\theta_q(\beta)=1-\theta_1\beta-\theta_2\beta^2-...-\theta_q\beta^q$

$$
\begin{aligned}
y_t&=(1-\theta_1\beta-\theta_2\beta^2-...-\theta_q\beta^q)\epsilon_t \\
y_t&=\theta_q(\beta)\epsilon_t=MA(q)
\end{aligned}
$$

Weights of moving average do not have to add up to one.

In moving average case, stationarity is almost guranteed since you're dealing with white noise, which is stationary. Less constraints for moving average. There are other constraints for a moving average model in order for it to work properly.

$$
\begin{aligned}
y_t&=\epsilon_t-\theta\epsilon_{t-1} \\
&=y_t+\theta\epsilon_{t-1}
\end{aligned}
$$

Invertibility of an AR model

Infinite series:

$$\epsilon_t=y_t+\theta y_{y-1}+\theta^2y_{t-2}+\theta^3y_{t-3}...$$

You can convert an AR to an MA, but you will have an explosion of parameters.

Choose AR or AR(P) + MA(q). They do better together.

ARMA (p, q)
ARIMA = p, d, q

```{r}
library(tseries)
```

```{r}
ts.plot(lynx)
```


```{r}
lambda <- BoxCox.lambda(lynx)
lambda

ts.plot(BoxCox(lynx, lambda))
```

```{r}
ar1 <- arima.sim(list(order = c(1, 0, 0), ar = 0.75), n = 100)
ts.plot(ar1)
```

```{r}
ma1 <- arima.sim(list(order = c(0, 0, 1), ma = 0.9), n = 100)
ts.plot(ma1)
```

If it is AR1 process, you want $\phi<1$.

If it is AR2 process, you want
$\phi_1+\phi_2<1$
$\phi_2-\phi_1<1$
$|\phi_2|<1$

```{r}
arma21 <-
  arima.sim(list(
    order = c(2, 0, 1),
    ar = c(0.9,-0.75),
    ma = 0.9
  ), n = 100)

ts.plot(arma21)
```

```{r}
model <- auto.arima(arma21)

plot(forecast(model, h = 10))
```

Workflow:

* Determination: what model will you use for your problem? HOLT winters, AR model, MR model, ARIMA model

* Specify the model: in case of ARIMA, specify p, d, q

* Estimation of parameters

* Evaluation

Try `Arima`. You can pass the model and data.



