---
title: "Lecture 2"
author: "Joshua Goldberg"
date: "`r format(Sys.time(), '%B, %d %Y')`"
output:
  html_document:
    df_print: paged
  pdf_document: default
editor_options:
  chunk_output_type: inline
always_allow_html: yes
---

```{r Global-options, include=FALSE}
knitr::opts_chunk$set(fig.width=6, fig.asp=0.618, fig.path='Figs/',
                      warning=FALSE, message=FALSE)
```

```{r Preamble, echo=FALSE}
# Enter package in p_load()
# If package is not installed, p_load() will install and load the package
if(!"pacman" %in% rownames(installed.packages())) {
  install.packages("pacman")
  }
pacman::p_load(tidyverse, ggthemes, here)

# Set default ggplot theme to tufte
theme_set(ggthemes::theme_tufte())
```

```{r Copy-files, echo=FALSE, eval=FALSE}
# Enter files to load to project directory in from = "~/Downloads/your_file_name_here"
file.copy(from = "~/Downloads/", to = here::here(), 
          overwrite = TRUE, recursive = FALSE, 
          copy.mode = TRUE)
```

# Making your time series stationary

First difference is a good time series to make it stationary.

Stationary has many more powerful tools than non-stationary.

Any process that has a unit root is stationary ??

# Forecasting

Holt Winters model

Use when there is seasonality or trend. If one is missing, put zero for the variable.

## Two types

Additive

Multiplicative

# Regression

Time series regression: you have to ensure the time series is station.

## WOLD decomposition

A stationary (weak until stated otherwise) time series $y_t$ can be expressed as a sum of: 2TS = deterministic + stochastic.

Any stationary time series you can represent it as a stochastic process with white noise.

$y_t=\sum^\infty_{i=0} \psi w_{t-i}+\eta_t$

$\eta_t$ is deterministic
$\w_{t-i}$ is white noise
$\psi$ are the deviations (it's a constant weight)

$\psi_0=1$ and $\sum^\infty_{i=1}<\infty$

What is the underlying data generation process when we observe a time series.

For all stationary processes it is a summation of white noise and deterministic (which can be zero).

You do not always need a non-zero for $\eta_t$.

$\mu$ will be deterministic in a time series.























