---
title: "Assignment 4: Unemployment and GDP"
author: "Joshua Goldberg"
date: "`r format(Sys.time(), '%B, %d %Y')`"
always_allow_html: yes
output:
  pdf_document: default
  github_document: 
editor_options: 
  chunk_output_type: inline
---

```{r Global-options, include=FALSE}
knitr::opts_chunk$set(fig.width=6, fig.asp=0.618, fig.path='Figs/',
                      warning=FALSE, message=FALSE)
```

```{r Preamble, echo=FALSE}
# Enter package in p_load()
# If package is not installed, p_load() will install and load the package
if(!"pacman" %in% rownames(installed.packages())) {
  install.packages("pacman")
  }
pacman::p_load(tidyverse, ggthemes, here, tsibble, lubridate, tsibble, zoo,
               tseries, rlang, broom, forecast, janitor, lubridate, readxl,
               fpp, xts, TSA)

options(stringsAsFactors = FALSE)

# Set default ggplot theme to tufte
theme_set(ggthemes::theme_tufte())
```

# Data/Objective

The daily data is from Illinois Dept of Transporation (IDOT) for I80E 1EXIT (the 2nd data column) - note each data point is an hourly count of the number of vehicles at a specific location on I80E.

Use the daily data for last 2 weeks of June 2013 to develop an ARIMA forecasting model.

Objective is to forecast the hourly counts for July 1.

The actual data file for July 1 is included for you to test your estimate.

```{r}
data_files <- data.frame(
  file_name = dir("traffic-flow/"),
  date = ymd(dir("traffic-flow/") %>% str_remove_all("^I-57-|.xls$"))
) %>% slice(-1)

test_files <- data.frame(
  file_name = dir("traffic-flow/"),
  date = ymd(dir("traffic-flow/") %>% str_remove_all("^I-57-|.xls$"))
) %>% slice(1)

extract_excel <- function(file_name, date) {
  readxl::read_xls(
    paste("traffic-flow/", file_name, sep = "/"),
    skip = 2,
    range = cell_cols("C:E")
  ) %>%
    slice(3:(nrow(.) - 2)) %>%
    select(Time, I80E) %>%
    janitor::clean_names() %>%
    mutate(date = date,
           date_time = as.POSIXct(paste(date, time), format = "%Y-%m-%d %H:%M",
                                  tz = Sys.timezone(location = TRUE)))
}

train_data <- pmap_df(data_files, extract_excel) %>%
  mutate_at(vars(i80e), as.numeric) %>%
  as_tsibble(index = "date_time")

test_data <- pmap_df(test_files, extract_excel) %>%
  mutate_at(vars(i80e), as.numeric) %>%
  as_tsibble(index = "date_time")
```

# Explore

```{r}
ggAcf(train_data$i80e)
ggPacf(train_data$i80e)
```

Augmented Dickey-Fuller Test

```{r}
adf.test(train_data$i80e)
```


# Modeling

## Part 1

Use `ARIMA(p, d, q)` model to forecast. Find the model returned by R `auto.arima()`. Change the values of p and q and determine the best model using AICc and BIC. Do AICc and BIC select the same model as the best model?

```{r}
train_auto <- auto.arima(train_data$i80e, seasonal = FALSE)
train_auto
```

Exploring more models.

```{r}
parameters <- list(mod2 = c(2, 0, 3), mod3 = c(3, 0, 3), mod4 =c(3, 0, 2), mod5 = c(3, 0, 1))
models <- map(parameters, ~ Arima(train_data$i80e, order = .x))
models
```

AIC and BIC select the same model.

```{r}
aiccs <- map_dbl(models, "aicc") %>% sort(decreasing = TRUE)
bics <- map_dbl(models, "bic") %>% sort(decreasing = TRUE)

glue::glue("AIC{2:5}: {aiccs}")
cat("\n")
glue::glue("BIC{2:5}: {bics}")
```

Examining residuals and normality for each models.

```{r}
walk(models, ~ checkresiduals(.x))
```

# Part 2

Use day of the week seasonal `ARIMA(p,d,q)(P,Q,D)s` model to forecast for July 1 (which is a Monday) - note use the hourly data.

```{r}
train_daily <- ts(train_data$i80e, start = c(16, 1), frequency = 24 * 7)
autoplot(train_daily)

model_day_of_week = auto.arima(train_daily, seasonal = TRUE)
summary(model_day_of_week)
```

Let's forecast the values for July 1st.

```{r}
forecast_day_of_week <- forecast(model_day_of_week, h = 24)
plot(forecast_day_of_week)
```


# Part 3

Use hour of the day seasonal `ARIMA(p,d,q)(P,D,Q)s` model to forecast for the hours 8:00, 9:00, 17:00 and 18:00 on July 1.

```{r}
train_hourly = ts(train_data$i80e, start = c(16, 1), frequency = 24)
autoplot(train_hourly)
```

```{r}
model_hourly = auto.arima(train_hourly, seasonal = TRUE)
summary(model_hourly)
```

```{r}
forecast_hourly <- forecast(model_hourly, h = 24)
plot(forecast_hourly)
```

# Part 4

For the July 1 8:00, 9:00, 17:00 and 18:00 forecasts, which model is better (part 2 or part 3)?

```{r}
test_data_ts <- ts(test_data[, 3], start = c(16, 1), frequency = 24)
```


```{r}
indexes <- c(8, 9, 17, 18)
test_data_ts[indexes]
forecast_day_of_week$mean[indexes]
forecast_hourly$mean[indexes]
```

```{r}
sse_weekly = sum((forecast_day_of_week$mean[indexes] - test_data_ts[indexes])^2)
sse_hourly = sum((forecast_hourly$mean[indexes] - test_data_ts[indexes])^2)

glue::glue("see_weekly: {sse_weekly}")
glue::glue("sse_hourly: {sse_hourly}")
```

Based on SSE, the weekly model performs better.




